### 2021-11-15 documents/sections/extensions.tex
tiriyoan *t- review

\subsection{\PTir \rc{t-}}
\label{sec:taranoan}
The Taranoan languages were already grouped together by \textcite{girard1971proto}, consisting of closely related \trio and \akuriyo (subsumed here under the moniker Tiriyoan), and the more distant member \carijo.
\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar.
He faced an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
\footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
\input{floats/ptirreg}
The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
In fact, even the issue of \emph{when} this happened is open.
It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation.
Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it taking place already in \PTar.}
%
Regarding \gl{s_a_} verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar; we offer our reconstructed \PTir forms.
To this list, the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}} can be added, which has first person \obj{w-}, at least in \trio.

>>>>

\subsection{\PTir \rc{t-}}
\label{sec:taranoan}
The moniker Tiriyoan subsumes \trio and \akuriyo, the more closely related of the three Taranoan languages already identified by \textcite{girard1971proto}, the more distant member being \carijo.
\textcite{meira1998proto} provides an extensive phonological, morphological, and lexical reconstruction of \PTar.
He faces an interesting puzzle in the \setone paradigms of \trio and \akuriyo: \PC \gl{1}>\gl{3} \rc{t-} and \gl{1}\gl{s_a_} \rc{w-} seem to have switched places.
This resulted in a regular \gl{1}\gl{s_a_} marker of the form \envr{\rc{t͡ʃ-}}{\obj{e}}, \envr{\rc{t-}}{\obj{ə}} \pcref{tab:ptirreg}.%
\footnote{The latter allomorph was subsequently replaced by \obj{k-} in \akuriyo \pcref{sec:akuriyo}.}
\input{floats/ptirreg}
The question of how this switch happened in detail \parencite[107--112]{meira1998proto} still has no answer, although it seems necessary to assume a scenario whereby both \rc{t-} and \rc{w-} for a time occurred on both transitive and intransitive verbs \parencite[112]{meira1998proto}.\footnote{
In fact, even the issue of \emph{when} this happened is open.
It could have happened at the \PTar stage, but the subsequent introduction of \obj{j-} in \carijo \pcref{sec:carijo} would have erased any traces of such an innovation.
Accordingly, \textcite{meira1998proto} hesitates to assign this extension to a specific proto-language.
We take a conservative stance and reconstruct it to \PTir only, but acknowledge the possibility of it taking place already in \PTar.}
%
Regarding \gl{s_a_} verbs unaffected by the spread of \rc{t-}, \textcite{meira1998proto} reconstructs the first four items in \cref{tab:ptirverbs} as taking \rc{w-} in \PTar -- attentive readers my recognize \trio \qu{to go} -- for which we provide our reconstructed \PTir forms.
To this list, the other copular root \rc{eʔi} \parentext{\PTar \rc{et͡ʃi} \parencite[165]{meira1998proto}} can be added, which has first person \obj{w-}, at least in \trio.
